Scalar and Representative Models for Polarimetric SAR Data

### 2013-05-17 paper_dev/article.tex
2013May_0517

%\documentclass[a4paper,10pt]{article}
\documentclass[journal,12pt,draftcls,onecolumn]{IEEEtran}
%\documentclass[a4paper, 10pt, conference]{ieeeconf}

@@

%\nonstopmode

%opening
\title{
  %Consistent and Homoskedastic Measures of Distance for Heteroskedastic POLSAR Data
  Scalar and Representative Models for Polarimetric SAR Data
}

%\author{Thanh-Hai Le, Ian McLoughlin}
\author{Thanh-Hai~Le,
        Ian~McLoughlin, 
	and Chan-Hua~Vun%
\thanks{Thanh-Hai~Le and Chan-Hua~Vun are with School of Computer Engineering, 
Nanyang Technological University, Singapore. Ian~McLoughlin is with School of Information Science and Technology,
University of Science and Technology of China.
}% <-this % stops a space
%\thanks{The authors wish to thank Dr. Ken-Yoong Lee and Dr. Timo Brestchneider of EADS Innovation-Works Singapore for 
%	their initial discussions and for providing us the RADAR-SAT2 imagery used in this paper. }% <-this % stops a space
\thanks{Manuscript received ?, 2013; revised ?.}}

@@

\begin{abstract}
  While statistical models are important in understanding the stochastic nature of Polarimetric SAR (POLSAR),
  scalar discrimination measures are routinely used in processing this multi-dimensional data.
  Ideally speaking, the latter should be drived from the former.
  As far as we know, however, it has not exactly been so for the case of POLSAR.
  In this paper, a couple of statistical models for the determinant of the POLSAR covariance matrix and its log-transformed derivations are proposed.
  These scalar models for the determinant of the POLSAR covariance matrix are shown to be respresentative
    since this observable is neatly transformed into the traditional SAR intensity
    when the generic multi-dimensional polarimetry is collapsed into the case of one-dimensional SAR. %case of traditional SAR data.
  Similar to SAR, the generic model for POLSAR is also shown to be multiplicative and heteroskedastic in the original domain 
    and the logarithmic transformation converts this into a more familiar additve and homoskedastic model.
  Moreover, the scalar property of these statistical models lead to the derivation of several statistical discrimination measures. %homoskedastic measures of distance.
  All of these statistical models are validated on real-life captured data
    and they are shown to be applicable even when the theoretical ``independent samples'' assumption is violated in practice.
  %The paper also illustrates a couple of benefits of these models, specifically for using the homoskedastic discrimination measures.
  %A couple of benefits for using these homoskedastic dis-similarity measures are then illustrated in this paper.
  %These homoskedastic measures of distance
  The combination of scalar, additive and homoskedastic properties of the derived statistical models bring about a couple of benefits, which are also described in this paper.
  First, compared to the original multiplicative domain, %dis-criminative measure of ratio,
    the additive log-transformed fit more naturally with the linear nature of digital imagery.
  Second, since the Gauss-Markov theorem is again applicable in the homoskedastic log-transformed domain,
    the use of Mean Squared Error (MSE) as an reliable objective function for POLSAR speckle filters is briefly explored.
%This paper presents a novel homoskedastic and additive model for polarimetric SAR (POLSAR) data. 
%The statistical properties of this new model  enable a number of different consistent measures of distance for the POLSAR covariance matrix. 
%The dissimilarity measurements, 
%computed from the determinant of the POLSAR covariance matrix, are initially developed for both partial (2$\times$2) and full (3$\times$3 monostatic case) polarimetric SAR data.
%It is shown that when the multi-polarization POLSAR is collapsed into the single-polarization SAR,
%  this covariance matrix determinant is neatly transformed into the traditional SAR intensity.
%This puts the SAR intensity statistics to lie within the natural coverage of this new theoretical model for POLSAR.
%Not only comprehensive theoretically, the statistical model is also robust in handling practical data. 
%It can be validated and match reasonably well  with practical data.
%Moreover, the match can be further improved 
%  if the model validation process includes estimation of the dataset's Effective Number of Looks (ENL) 
%  instead of blind acceptance of the number of looks figure reported by the SAR processor.
%The application of these dissimilarity measures is demonstrated in this paper for evaluating the effectiveness of POLSAR speckle filters.  
\end{abstract}

\begin{IEEEkeywords}
Polarimetric Synthetic aperture radar, Electromagnetic Modeling, Multidimensional signal processing  
\end{IEEEkeywords}

@@

\section{Introduction}
%\section{POLSAR data: statistical analysis and logarithmic transformation}

In the past decades, the exponential growth in computational power has made the once overly-excessive computationaly-demanding SAR technology now become a feasible and preferred earth observation solution.
As state-of-the-art technology, the SAR technique has been extended in a few directions, one of which is the polarimetric SAR or POLSAR.
POLSAR is the natural extension from SAR exploiting the natural polarization property of Electro-Magnetic (EM) waves.
Similar to the extension from black-and-white images to color photography, the polarimetric extension brought about the multi-channels POLSAR data as compared to the traditional one-channel SAR data.

With (POL)SAR data becomes cheaper and more available, the recent research emphasis is on understanding and developing applications for the data. %it is of great important to understand the data.
The complex data, however is stochastic by nature, due to the interference phenomena of EM waves. %, SAR measurements are stochastic by nature.
Under this condition, it is very important to understand the statistical models for (POL)SAR data.

While the statistical models for homogeneous areas of one-dimensional SAR data have been developed,
%  in extending the model toward heterogeneous images, the impact of heteroskedasticity has virtually been ignored.
%Specifically the concept of distance as a measure of dis-similarity is seriously flawed, as it is widely known that ratio offers a better discrimination measure TODOCITE.
%This thesis starts off by proposing the logarithmic transformation which is shown converting the SAR data into an homoskedastic model.
%From this log-transformed model, a few consistent measures of distance is proposed. 
%in a sense the distance concept is re-emerged, as the log-transformation converts the widely used ratio into standard subtraction, and thus distance.
  in extending our understanding towards multi-dimensional POLSAR data, an important issue needs to be addressed.
The trouble that the high dimensional data bring about is that there exists not one, as the intensity in the one-channel SAR, but many observables quantities in multi-channel POLSAR.
While different statistical models have been developed for different POLSAR observables,
  for a statistical model to be useful, scalar discrimination measures need to be derived from it.
Thus practical application for POLSAR data processing requires the dis-similarity measure to be scalar, consistent and preferably homoskedastic on one hand.
On the other hand, the observable quantity being modelled needs to be naturally representative for the high dimensional POLSAR data.

In this paper, the determinant of the POLSAR covariance matrix is proposed as an observable quantity to study.
The representative power of this observable is demonstrated when the multi-dimensional POLSAR data is collapsed into the traditional one-dimensional SAR scenario.
Under this transformation, the determinant of the POLSAR covariance matrix is converted into the standard SAR intensity.
Statistical models for this heteroskedastic POLSAR determinant and homoskedastic log-determinant are then derived.
And since POLSAR can be viewed as a multi-dimensional extension of SAR, the paper describes how the standard statistical models for SAR can be put within the natural coverage of this generic models for POLSAR.

%There are a few benefits of using the homoskedastic measures of distance for POLSAR.
%Firstly, compared to the multiplicative dis-similarity measure of ratio, these additive measures of distance fit more naturally with the linear scale nature of digital imagery.
%Secondly, within the homoskedastic log-transformed domain, where the distance concept as well as the Gauss Markov theorem is again applicable, the universal MSE deserves to be reviewed as a reliable objective function / criteria for (POL)SAR estimators.
%This may open up a number of different research directions for future studies of (POL)SAR.
%As example demonstrations of how these models and distance measures can be used, a novel clustering algorithm, a new speckle filter and various procedures in using MSE to evaluate speckle filters are also presented as part of the thesis.

Specifically, the rest of this paper is structured as follows.
After the related work in literature is critically reviewed in Section \ref{sec:lit_review},
  Section \ref{sec:theoretical_model} presents the theoretical models for both partial and full polarimetric SAR and their derived discrimination measures.
The representative power of the scalar models are illuminated in Section \ref{sec:sar_special_case_of_polsar},
  which details how the proposed models for the multi-variate POLSAR also includes the well-known models for the traditional SAR as its special univariate case.
%can be used to derive the well-known models for the univariate SAR data as its special case.
Besides being theoretically comprehensive, 
  the model is also shown, in Section \ref{sec:polsar_models_validation}, %in section \ref{sec:improve_the_match_bw_theory_practice},
  to be robust in handling practical data. % imperfection that violate conventional assumptions. 
To demonstrate the application of these scalar and representative models, 
  Section \ref{sec:evaluating_polsar_filters} explores how they can be used to evaluate POLSAR speckle filters.   
In the end, Section \ref{sec:discussion_conclusion} provides the final conclusion of the paper.
%critically reviews and discusses related literature  
%  before the paper is concluded in Section \ref{sec:conclusion}.

%In this paper, the above results are exploited in the context of POLSAR data processing. In particular, a few scalar, consistent and homoskedastic measures of distance are proposed. The paper is structured as follows:
%  after the preceding introduction to the basics of POLSAR statistical analysis,
%  the second section briefly reviews related work in published literature.
%Section \ref{sec:polsar_heterosked_model_and_log_transform} presents the theoretical models together with conclusive evidence for two related points.
%The first point is that: POLSAR data is multiplicative and heteroskedastic in its original domain.
%And the second conclusion being: log-transformation converts it into an additive and homoskedastic model.
%Consequently a few consistent measures of distance are presented in section \ref{sec:distance_measure}.
%After the model is presented for both partial and full polarimetric SAR data,
%  section \ref{sec:sar_special_case_of_polsar} discusses
%    how the proposed models for the multi-variate POLSAR can be used to derive the well-known models for the univariate SAR data as its special case.
%Besides being theoretically comprehensive, 
%  the model is also shown, in Section \ref{sec:polsar_models_validation}, %in section \ref{sec:improve_the_match_bw_theory_practice},
%  to be robust in handling practical data. % imperfection that violate conventional assumptions. 
%To demonstrate the application of the consistent and homoskedastic dis-similarity measures, 
%  Section \ref{sec:evaluating_polsar_filters} explores how they are used for evaluating POLSAR speckle filters.   
%Finally, Section \ref{sec:discussion_conclusion} critically reviews and discusses related literature  
%  before the paper is concluded in Section \ref{sec:conclusion}.

@@

This section critically review the related work in published literature.
Specifically, the first sub-section shows that while many discrimination measures have been proposed for POLSAR data,
  almost all of them depends on the likelihood statistical test for complex Wishart distribution.
Unfortunately, in the original proposed work \cite{Conradsen_2003_TGRS_4}, while an exact distribution is to be expected, only an asymptotic distribution was given.
The second sub-section reviews various statistical models for different POLSAR observables.
Its purpose is to show that none of these models, as far as we know, led to statistically consistent discrimination measures.

\subsection{POLSAR Discrimination Measures}

This sub-section reviews some relevant and published works relating to different measures of distance for POLSAR data. 
In particular, a few different matrix distances have been proposed and evaluated in recent review papers\cite{Dabboor_2013_IJRS_1492}\cite{Kersten_2005_TGRS_519}
will be briefly reviewed.
%specifically a few different matrix distances have been proposed and evaluated in recent review papers\cite{Dabboor_2013_IJRS_1492}\cite{Kersten_2005_TGRS_519}

The commonly used measure of distance for matrices are either the Euclidean or the Manhattan distance,
  which are defined in the following equations, respectively:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $A_{i,j}$ denotes the (i,j) elements of the matrix A,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR covariance matrix, these dis-similarity measures are not widely used 
  probably because of the multiplicative nature of the data.
%In the context of polsar covariance matrix, however, they are not widely used 
%  probably because of the multiplicative nature of the data.

In the field of POLSAR, the Wishart Distance is probably the most widely used as part of the well-known Wishart Classifier \cite{Lee_1999_TGRS}.
The distance is defined as \cite{Lee_1994_IJRS_2299}
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation}
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested a number of other dissimilarity measures.
These include the asymmetric and symmetric refined wishart distance \cite{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \cite{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \cite{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \cite{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}
.

In comparison to these formulae, the distance measures for two covariance matrices proposed in this paper can be written as:
\begin{equation}
  d(C_x,C_y) =  \ln|C_x| - \ln|C_y| 
\end{equation}
At first glance, while the propose formula may be a lot simpler, it still operates in a similar way to others in making use of the log-determinant.

Closer investigation of these dis-similarity measures reveals that most of them are related to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered as the special case of the Wishart Statistical Test distance, where the two data set have the same number of look, i.e. $L_x=L_y$.
It is also intuitively trivial to arrive at the conclusion that the contrast measure of distance proposed in this paper has fixed statistical behaviour, by comparing the symmetric and asymmetric versions of the refined Wishart distance, assuming that they are shown to follow fixed distributions.

The close relation among these is further supported by the fact that
all of the papers proposing these measures referenced the statistical model developed in \cite{Conradsen_2003_TGRS_4} as their foundation.
%Interestingly, the change detection model in \cite{Conradsen_2003_TGRS_4} also make use of determinant ratio and log determinant in its derivation.  
%The next sub-section will illustrate how the log-determinant model presented in this paper can offer an alternative and simple derivation for the change-detection statistics.
In \cite{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrix $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

@@

To detect changes, a test statistics is developed based on this measure of distance.
This means a distribution is to be derived for the dissimilarity measure.
However, originally in the proposed work \cite{Conradsen_2003_TGRS_4}, only an asymptotic distribution is derived.
%In this sub-section, a statistical model is developed for this measure of distance.
Since both $Z_x$ and $Z_y$ follow a complex Wishart distribution with $L_x$ and $L_y$ degrees of freedom,
  $Z_x+Z_y$ also follows the complex Wishart distribution with $(L_x + L_y)$ degrees of freedom.
In view of the models for $|C_v|$ and $\ln|C_v|$ developed in this paper (Eqns. \ref{eqn:determinant_distribution} \& \ref{eqn:log_determinant_distribution}),
  it is evident that not only the bound for $\ln Q$, or equivalently $Q$, can be derived
  but the whole statistical distribution for it can be simulated as well.

\subsection{Scalar Observables and Statistical Models for POLSAR Data}

Of course, discrimination measures are not the only way to carry out classification tasks in POLSAR.
Another common approach is to carry out classification using POLSAR observables identified by different target decomposition theorems.
In \cite{Alberga_2008_IJRS_4129}, the performance of different scalar POLSAR observables for classification purposes is evaluated.
While many scalar observables for POLSAR were presented, their corresponding statistical models and classifiers were not available.
Furthermore, at its conclusion, the paper indicated that ``it is impossible to indentify the best one''
  as these ``representations ... do not show a precise trend''.
%Instead an MLP classification approach were employed. 
To be fair, these observables are identified to describe a decomposed portion of the complex POLSAR data,
  and thus they are not to be used individually for the purpose of single-handedly representing POLSAR data.

Given the joint distribution for POLSAR is known to be the multi-variate Complex Wishart distribution,
  it is possible to derive the scalar statistical models for some univariate POLSAR observables.
However, such derivations are no trivial tasks,
  as the data is much more complex and the field is much less matured than the traditional SAR statistical modelling.
Thus the strategy of the pioneering work is to tackle simpler versions of polarimetry.

POLSAR data is commonly represented by its covariance matrix.
For full-pol the covariance matrix has the form of
\begin{equation*}
C_3 = 
\begin{bmatrix}
S_{hh}^2 & \sqrt{2}S_{hh}S_{hv}^* & S_{hh}S_{vv}^* \\ 
\sqrt{2}S_{hh}^*S_{hv} &2S_{hv}^2  & \sqrt{2}S_{hv}S_{vv}^*\\ 
S_{hh}^*S_{vv} & \sqrt{2}S_{hv}^*S_{vv} & S_{vv}^2
\end{bmatrix}
\end{equation*}
For part-pol, it reduces to either:
\begin{equation*}
  C_{2,h} =
  \begin{bmatrix}
    S_{hh}^2 & S_{hh}S_{hv}^* \\
    S_{hh}^*S_{hv} & S_{hv}^2
  \end{bmatrix};
  C_{2,v} =
  \begin{bmatrix}
    S_{hv}^2 & S_{hv}S_{vv}^* \\
    S_{hv}^*S_{vv} & S_{vv}^2
  \end{bmatrix}
\end{equation*}
Obviously, the statistical behavior of $C_3$ can be understood by understanding the behavior of $C_{2h},C_{2v}$ and
\begin{equation*}
  C_{2,c} = 
\begin{bmatrix}
    S_{hh}^2 & S_{hh}S_{vv}^* \\
    S_{hh}^*S_{vv} & S_{vv}^2
\end{bmatrix}
\end{equation*}

While $C_{2v}, C_{2h}$ are physically related to the Stokes parameters of EM waves,
  $C_{2c}$ is a totally made-up mathematical term from the full-pol covariance matrix.
In fact, in the field of Polarimetric SAR, $C_{2v}, C_{2h}$ is called cross-pol and $C_{2c}$ is called co-pol covariance matrices.
Their main difference lies in investigating the so-called complex correlation coefficient, defined as:
\begin{equation*}
  \rho = \frac{E \left( S_{pq}S_{rs}^* \right) }{\sqrt{E \left( S^2_{pq} \right) E \left( S^2_{rs} \right) }} = |\rho| e^{j \phi}
\end{equation*}
with $\phi=arg(\rho)$ is the phase difference angle.
In $C_{2v}, C_{2h}$ of optical physics, $|\rho_v|,|\rho_h|$ are insignificant,
  and $\phi_v,\phi_h$ are very random, due to little correlation between orthogonal polarization in monochromatic EM waves.
%By analogy, in POLSAR ``natural surfaces'' data, it is common to consider: $|\rho_v|=|\rho_h|=0$
Their equivalent lie in the ``natural surface'' case of POLSAR,
  where  it is common to consider: $|\rho_v|=|\rho_h|=0$.
This is, however, not true in the general case of POLSAR.  
Specifically for $C_{2c}$, since there are correlation between the two co-pol channels,
  $|\rho_c|$ is almost always significant and $\phi_c$ is also not entirely random.

In the field of optical physics,
  under the assumption of no correlation between cross-polarized channels,
  \cite{Barakat_1985_IJOptics}, \cite{Eliyahu_1993_PhysRevE_2881} and \cite{Brosseau_1995_AppliedOptics_4788} derived the statistical models for $C_{2h},C_{2v}$.
The observables that have been studied are:
\begin{enumerate}
\item the polarization ellipse parameters \cite{Barakat_1985_IJOptics}: i.e.
  \begin{enumerate}
  \item the total power $S_0=C_{11}^2 + C_{22}^2$,
  \item the ratio of minor to major axis of the polarization ellipse  $\epsilon=\lambda_2/\lambda_1$,
  \item and $\psi$ the angle that the ellipse major axis makes with horizontal axis,
  \end{enumerate}
\item the Stokes parameters \cite{Eliyahu_1993_PhysRevE_2881}: i.e.
  \begin{enumerate}
  \item $S_0 = C_{11}^2 + C_{22}^2$
  \item $S_1 = C_{11}^2 - C_{22}^2$
  \item $S_2 = \Re (C_{12})$
  \item $S_3 = \Im (C_{12})$
  \end{enumerate}
\item and the normalized Stokes parameters $s_i=S_i/S_0, i>0$ \cite{Brosseau_1995_AppliedOptics_4788}
\end{enumerate}
In the field of POLSAR, under the assumption of significant correlation between polarized channels,
  \cite{Joughin_1994_TGRS_562}, \cite{Lee_1994_TGRS_1017} and \cite{Touzi_1996_TGRS_519} derived the statistical models for note only $C_{2h},C_{2v}$ but also $C_{2c}$.
Specifically the following observables were studied:  
  \begin{enumerate}
  \item cross-pol ratio $r_{HV/HH} = |S_{HV}|^2/|S_{HH}|^2$ \cite{Joughin_1994_TGRS_562},
  \item co-pol ratio $r_{VV/HH} = |S_{VV}|^2/|S_{HH}|^2$ \cite{Joughin_1994_TGRS_562},
  \item co-pol phase difference $\phi_{VV/HH} = arg(S_{VV}S_{HH}^*) $ \cite{Joughin_1994_TGRS_562} \cite{Lee_1994_TGRS_1017},
  \item magnitude $g=|avg(S_{pq}S_{rs}^*)|$ \cite{Lee_1994_TGRS_1017},
  \item normalized magnitude $\xi = \frac{|avg(S_{pq}S_{rs}^*)|}{\sqrt{avg(|S_{pq}|^2) avg(|S_{rs}|^2)}}$ \cite{Lee_1994_TGRS_1017},
  \item intensity ratio $w = avg(|S_{pq}|^2)/avg(|S_{rs}|^2)$ \cite{Lee_1994_TGRS_1017},
  \item and the Stokes parameters $S_i,0 \leq i \leq 3$ \cite{Touzi_1996_TGRS_519}. 
  \end{enumerate}

Furthermore, other studies has offered
  %\cite{Lopez-Martinez_2003_TGRS_2232} and \cite{Erten_2012_Sensors_2766} gives
  the statistical model for
  each element of the POLSAR covariance matrix $S_{pq}S_{rs}^*$ \cite{Lopez-Martinez_2003_TGRS_2232}
  as well as the largest eigen-value of the covariance matrix $\lambda_1$ \cite{Erten_2012_Sensors_2766}.
While these models undoubtedly help in understanding the SAR data,
  individually none of them statisfies the dual criteria of
  1) resulting in statistically consistent discrimination measures and
  2) being an representative observable for the complex POLSAR data.
  %that they model have the power to represent the multi-demensional POLSAR covariance matrix.

\section{The Scalar Statistical Models for POLSAR}  
\label{sec:theoretical_model}

In this section, several scalar statistical models for POLSAR are presented.
After the first sub-section briefly reviews related work in published literature,
  sub-section \ref{sec:polsar_heterosked_model_and_log_transform} presents the theoretical models together with conclusive evidence for two related points.
The first point is that: POLSAR data is multiplicative and heteroskedastic in its original domain.
And the second conclusion being: log-transformation converts it into an additive and homoskedastic model.
Consequently a few consistent measures of distance are presented in section \ref{sec:distance_measure}.

\subsection{The Basics of POLSAR Statistical Analysis}

>>>>

\documentclass[journal,12pt,draftcls,onecolumn]{IEEEtran}
%\documentclass[a4paper, 10pt, conference]{ieeeconf}

@@

%opening
\title{
  Scalar and Representative Models for Polarimetric SAR Data
}

%\author{Thanh-Hai Le, Ian McLoughlin}
\author{Thanh-Hai~Le,
        Ian~McLoughlin, 
	and Chan-Hua~Vun%
\thanks{Thanh-Hai~Le and Chan-Hua~Vun are with School of Computer Engineering, 
Nanyang Technological University, Singapore. Ian~McLoughlin is with School of Information Science and Technology,
University of Science and Technology of China.
}% <-this % stops a space
%\thanks{The authors wish to thank Dr. Ken-Yoong Lee and Dr. Timo Brestchneider of EADS Innovation-Works Singapore for 
%	providing us the RADAR-SAT2 imagery used in this paper. }% <-this % stops a space
\thanks{Manuscript received ?, 2013; revised ?.}}

@@

\begin{abstract}
This paper presents several novel scalar statistical models to represent the determinant of both partial and full polarimetric SAR (POLSAR) covariance matrices.
Compared to other scalar statistical models for POLSAR,
  the proposed models are highly representative of the multi-dimensional data
  and enable useful discrimination measures to be determined easily. %consequently they lead to useful discrimination measures.
When the multi-dimensional models are collapsed to a one-dimensional case,
  they are shown to exactly match the traditional statistical models for SAR intensity.
The paper validates the theoretical models together with their derived discrimination measures using real world POLSAR data.  
The proposed generic models are envisioned to enable the adaptation of many existing SAR data processing techniques for use with POLSAR data.
An  example is also shown where the determinant-ratio and its additive version, contrast, are applied in the evaluation of POLSAR speckle filters.
\end{abstract}

\begin{IEEEkeywords}
Polarimetric Synthetic Aperture Radar, Electromagnetic Modeling, Multidimensional Signal Processing  
\end{IEEEkeywords}

@@

\section{Introduction}

During the past decades, exponential growth in computing power has allowed once computationally-demanding Synthetic Aperture Radar (SAR)
technology to become a feasible and preferred technique for earth observation applications.
SAR technology has since extended in a few directions, one of which is polarimetric SAR (POLSAR).
POLSAR extends SAR by exploiting the natural polarization property of Electro-Magnetic (EM) waves,
  leading to the availability of multi-channel POLSAR data, compared to traditional one-channel SAR data.

While statistical models are important in understanding the stochastic nature of both SAR and POLSAR, in extending our understanding towards multi-dimensional POLSAR, an important issue needs to be addressed.
Namely that there now exists not one but many observables quantities in multi-channel POLSAR (as opposed to single-channel SAR intensity).
The research community has responded largely by developing different statistical models for various POLSAR observables.
However for a statistical model to be useful, scalar discrimination measures generally need to be derivable from it.
Thus, practical applications for POLSAR data processing require dis-similarity measures that are scalar, consistent and preferably homoskedastic on the one hand.
On the other hand, the observable quantity being modelled should be naturally representative of the high dimensionality POLSAR data.

This paper presents novel statistical models for multi-dimensional POLSAR data which generalize the commonly used statistical model for the one-dimensional SAR data and thus enables the adaptation of many existing SAR data processing techniques for POLSAR data.
Specifically statistical models for the heteroskedastic POLSAR determinant and its homoskedastic log-domian determinant, together with several other discrimination measures are derived in section  \ref{sec:theoretical_model}.
The representative properties of these observables are then demonstrated in section \ref{sec:sar_special_case_of_polsar}
  where the multi-dimensional POLSAR data is collapsed to a one-dimensional SAR scenario.
With this transformation, the determinant of the POLSAR covariance matrix is shown to be equivalent to SAR intensity,
  which also brings many standard statistical models used with SAR  under the umbrella of these generic POLAR models.
The proposed theoretical models are then validated against real life practical data in section \ref{sec:polsar_models_validation}.
Section \ref{sec:discussion} briefly describes how evaluation of an existing SAR data processing technique, the speckle filter, can be extended to  POLSAR data processing using this approach.
This is followed by a conclusion in Section \ref{sec:conclusion}.

@@

This section outlines related published work.
In particular, section \ref{sec:lit_models} reviews various scalar statistical models used for different POLSAR observables.
Its purpose is to show that none of those  models have led to statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthen these findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
However, while an exact statistical distribution is needed for the test,
  only an symptotic distribution is used in the original work \cite{Conradsen_2003_TGRS_4}.

\subsection{Investigated Scalar Observables for POLSAR Data and their Statistical Models}
\label{sec:lit_models}

Different target decomposition theorems have identified many different scalar observables for complex POLSAR data.
In \cite{Alberga_2008_IJRS_4129}, the performance of different scalar POLSAR observables is evaluated for classification purposes.
While many scalar observables for POLSAR were presented, their corresponding statistical models and classifiers were not available.
Furthermore, at its conclusion, the paper indicated that: it is impossible to identify the best representation.
To be fair, those observables were identified to describe a decomposed portion of the complex POLSAR data,
  rather than providing a single unified representation of the POLSAR data.

Given that the joint distribution for POLSAR is known to be the multi-variate complex Wishart distribution,
  it is possible to derive the scalar statistical models for some univariate POLSAR observables.
This becomes an alternative approach used in the study of POLSAR data.
However, such derivations are no trivial task, and so far, only a handful of such statistical models have been proposed, including the following:
  \begin{enumerate}
  \item cross-pol ratio $r_{HV/HH} = |S_{HV}|^2/|S_{HH}|^2$ \cite{Joughin_1994_TGRS_562},
  \item co-pol ratio $r_{VV/HH} = |S_{VV}|^2/|S_{HH}|^2$ \cite{Joughin_1994_TGRS_562},
  \item co-pol phase difference $\phi_{VV/HH} = arg(S_{VV}S_{HH}^*) $ \cite{Joughin_1994_TGRS_562} \cite{Lee_1994_TGRS_1017},
  \item magnitude $g=|avg(S_{pq}S_{rs}^*)|$ \cite{Lee_1994_TGRS_1017},
  \item normalized magnitude $\xi = \frac{|avg(S_{pq}S_{rs}^*)|}{\sqrt{avg(|S_{pq}|^2) avg(|S_{rs}|^2)}}$ \cite{Lee_1994_TGRS_1017},
  \item intensity ratio $w = avg(|S_{pq}|^2)/avg(|S_{rs}|^2)$ \cite{Lee_1994_TGRS_1017},
  \item and the Stokes parameters $S_i,0 \leq i \leq 3$ \cite{Touzi_1996_TGRS_519}. 
  \end{enumerate}

More recently, statistical models for
  each element of the POLSAR covariance matrix $S_{pq}S_{rs}^*$ \cite{Lopez-Martinez_2003_TGRS_2232}
  as well as for the largest eigen-value of the covariance matrix $\lambda_1$ \cite{Erten_2012_Sensors_2766} have been studied.
While these models undoubtedly help in understanding POLSAR data,
  none of these observables have been shown to meet the dual criteria of
  1) resulting in statistically consistent discrimination measures and
  2) being representative of the complex POLSAR data.

\subsection{POLSAR Discrimination Measures}
\label{sec:lit_measures}

The commonly used measure of distance for matrices are either the Euclidean or Manhattan distances, defined as:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $A_{i,j}$ denotes the (i,j) elements of the matrix A,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR covariance matrix, these dis-similarity measures are not widely used 
  probably because of the multiplicative nature of the noisy data.

In the field of POLSAR, the Wishart distance is probably the most widely used as part of the well-known Wishart classifier \cite{Lee_1999_TGRS}.
It is defined as \cite{Lee_1994_IJRS_2299}
\begin{equation}
  d(C_x,C_y) = \ln|C_y| + tr(C_xC_y^{-1})
\end{equation}
As a measure of distance, its main disadvantage is that $d(C_y,C_y) = \ln|C_y| \neq 0$.

Recent works have suggested a number of other dissimilarity measures.
These include the asymmetric and symmetric refined Wishart distance \cite{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \cite{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \cite{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \cite{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}
.

Closer investigation of these dis-similarity measures reveals that most are related in some way. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among the measures is further supported by the fact that
  all of their proposing papers referenced the same statistical model developed in \cite{Conradsen_2003_TGRS_4} as a foundation.
In \cite{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrix $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

@@

To detect changes, a test statistic is developed based on this measure of distance.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \cite{Conradsen_2003_TGRS_4}, only an asymptotic distribution has been proposed.

\section{Scalar Statistical Models for POLSAR}  
\label{sec:theoretical_model}

In this section, several scalar statistical models for POLSAR are presented.
After section \ref{sec:model_basic} briefly introduce the basic foundations, concepts and notation of this paper, %reviews related work in published literature,
  section \ref{sec:polsar_heterosked_model_and_log_transform}
  presents the theoretical models together with mathematical evidence for two related points.
The first point is that: the POLSAR data is multiplicative and heteroskedastic in its original domain.
And the second conclusion states: log-transformation converts it into an additive and homoskedastic model.
Several consistent measures of distance are then derived and presented in section \ref{sec:distance_measure}.

\subsection{Basics of POLSAR Statistical Analysis}
\label{sec:model_basic}
